Innovations Report: Procedural Fur System

### 2018-03-18 docs/paper.tex
more on docs

\section{Introduction} \label{sec:introduction}
Introduction Text.
\subsection{Related work} \label{sec:related}

nvidia hairworks

\subsection{Existing solutions} \label{sec:existing}
Whilst many larger companies opt to develop their own software for handling fur, such as MPC's Furtility~\cite{furtility}, there are a number of commercially available solutions that are also widely used in production.

@@

\subsection{GPU acceleration} \label{sec:existingGPU}
Whilst it is not clear to what extent each of these solutions make use of GPU acceleration, they all perform fast enough to allow the user to manipulate them interactively. I therefore decided to monitor the CPU and GPU usage using \verb|htop| and \verb|nvidia-smi| respectively, before and after adding XGen content to a simple maya scene, to determine where the majority of data was being computed.

After applying the default XGen description onto a simple sphere with 760 triangles, there was no noticeable increase in both GPU utilisation and GPU memory usage. Increasing the density of the primitives from 1.0 to 10000.0 saw an increase of VRAM usage of roughly 350MiB, which I would attribute to storing the primitives in a buffer on the GPU, but there was no discernable increase in GPU utilisation. Increasing the density further to 10,000 saw a noticable spike in CPU usage, hitting 100 percent across two threads, leading me to believe that the primitive distribution computation is being done on the CPU.

However, when applying modifiers to the XGen description I noticed a small increase of roughly 2MiB in VRAM usage for the first modifier, yet no further increases for modifiers added after that. As this increase is quite small I would attribute it to storing the modifier attributes on the GPU, though I still noticed spikes in CPU usage when recalculating the XGen description once the modifier was added, which makes it inconclusive as to whether XGen uses any kind of GPU acceleration.

@@

\begin{figure}[htbp]\centering
\begin{center}
\begin{tabular}{||c||c|c|c||}
\hline
Mesh Triangle Count & Primitive Count & CPU Time (ms) & GPU Time (ms)\\
\hline
\hline
16 (Triangle Strip) & 10,000 & 2 - 5 & 1 - 2\\
\hline
 - & 100,000 & 18 - 41 & 8 - 24\\
\hline
 - & 1,000,000 & 120 - 150 & 77 - 110\\
\hline
4968 (Stanford Bunny) & 10,000 & 173 - 211 & 1 - 2\\
\hline
 - & 100,000 & 1718 - 1756 & 12 - 30\\
\hline
 - & 1,000,000 & 17218 - 17259 & 82 - 112\\
\hline
\end{tabular}
\caption{\label{figure:distributionPerformance} Performance metrics showing time taken to distribute varying numbers of primitives, on meshes of varying triangle counts. An Intel Core i7-6700 CPU and an NVIDIA GTX1080 GPU were used for testing.}
\end{center}
\end{figure}

@@

Whilst not necessary for usage in a broader pipeline, implementing a physically plausible hair shading model such as those described in \cite{Marschner}, would greatly improve the visual accuracy of my viewport renderings.

>>>>

\section{Introduction} \label{sec:introduction}
Generating and rendering hair and fur content for use in visual effects and games is typically a very computationally heavy task, due to the inherent complexity of handling thousands if not millions of geometric strands. Fortunately, modern hardware provides flexibility in regards to accelerating such tasks, which we are able to leverage though the use of hardware acceleration APIs such as OpenCL and CUDA.

\subsection{Related work} \label{sec:related}
Whilst many larger companies opt to develop their own software for handling fur, such as MPC's Furtility~\cite{furtility}, there are a number of commercially available solutions that are also widely used in production.

@@

\subsection{GPU acceleration} \label{sec:existingGPU}
Whilst it is not clear to what extent each of these solutions make use of GPU acceleration, they all perform fast enough to allow the user to manipulate them interactively. However, software such as the HairWorks \cite{hairworks} API and TressFX \cite{tressfx} library are developed primarily for use in games, so they are designed to utilise the GPU as much as possible. These solutions are able to perform both simulation and rendering of hair in real-time, providing interactivity that exceeds that of comparable software designed for VFX usage. It is worth noting that these tools do not provide a method for authoring hair and fur content, but instead designed to use load content generated by 3rd party authoring tools.

\begin{figure}[htbp]\centering
\includegraphics[width=1.0\linewidth]{images/hairworks}
\caption{\label{figure:hairworks} NVIDIA's HairWorks API is used in a handful of modern games, such as \textit{The Witcher 3} \protect\cite{witcher}, to provide realistic and interactive hair rendering.}
\end{figure}

% I therefore decided to monitor the CPU and GPU usage using \verb|htop| and \verb|nvidia-smi| respectively, before and after adding XGen content to a simple maya scene, to determine where the majority of data was being computed.

% After applying the default XGen description onto a simple sphere with 760 triangles, there was no noticeable increase in both GPU utilisation and GPU memory usage. Increasing the density of the primitives from 1.0 to 10000.0 saw an increase of VRAM usage of roughly 350MiB, which I would attribute to storing the primitives in a buffer on the GPU, but there was no discernable increase in GPU utilisation. Increasing the density further to 10,000 saw a noticable spike in CPU usage, hitting 100 percent across two threads, leading me to believe that the primitive distribution computation is being done on the CPU.

% However, when applying modifiers to the XGen description I noticed a small increase of roughly 2MiB in VRAM usage for the first modifier, yet no further increases for modifiers added after that. As this increase is quite small I would attribute it to storing the modifier attributes on the GPU, though I still noticed spikes in CPU usage when recalculating the XGen description once the modifier was added, which makes it inconclusive as to whether XGen uses any kind of GPU acceleration.

@@

\begin{figure}[htbp]\centering
\begin{center}
\begin{tabular}{||c||c|c|c||}
\hline
Mesh Triangle Count & Primitive Count & CPU (ms) & GPU (ms)\\
\hline
\hline
16 (Triangle Strip) & 10,000 & 2 - 5 & 1 - 2\\
\hline
 - & 100,000 & 18 - 41 & 8 - 24\\
\hline
 - & 1,000,000 & 120 - 150 & 77 - 110\\
\hline
4968 (Stanford Bunny) & 10,000 & 173 - 211 & 1 - 2\\
\hline
 - & 100,000 & 1718 - 1756 & 12 - 30\\
\hline
 - & 1,000,000 & 17218 - 17259 & 82 - 112\\
\hline
\end{tabular}
\caption{\label{figure:distributionPerformance} Performance metrics comparing time taken to distribute varying numbers of primitives, on meshes of varying densities, using the CPU and GPU. An Intel Core i7-6700 CPU and an NVIDIA GTX1080 GPU were used for testing.}
\end{center}
\end{figure}

@@

Whilst not necessary for usage in a offline pipeline, implementing a physically plausible hair shading model based on the techniques described in \cite{Marschner}, would greatly improve the visual accuracy of my viewport renderings.
